Causal Impact of Masks, Policies, Behavior on Early Covid-19 Pandemic in the U.S.

### 2020-09-07 tex/covidEffectOfNPI-revision.tex
update

Our estimates imply that nationally implementing mandatory face masks
for employees in public businesses on March 14th would have reduced the
growth rate of cases and that of deaths by approximately 10 percentage
points in late April. As shown in Figure \ref{fig:US-mask-dgrowth},
this leads to reductions of $21$\% and $34$\% in cumulative reported
cases and deaths, respectively, by the end of May with 90 percent
confidence intervals of $[9,32]$\% and $[19,47]$\%, which roughly
implies that 34 thousand lives could have been saved.\footnote{These estimates
are slightly lower than those reported in earlier releases of the paper on ArXiv.
One of the reason is that we chose to focus on a simplified version of the model 
as our base model, by averaging business closure policies as measure of shutdown. This produced more stable results when
conducting sensitivity results, in particular using fixed effects estimators.  
We also present several variants of base model, including
one with separate shutdown policies, in which the effects are slightly larger, in line
with the previous findings, but we focus reporting counterfactuals only 
for the base model. } 
This finding is
significant: given this potentially large benefit of reducing the
spread of Covid-19, mandating masks is an attractive policy instrument
especially because it involves relatively little economic
disruption. These estimates contribute to the ongoing efforts towards
designing approaches to minimize risks from reopening
\citep{stock2020b}.

@@

With exclusion restrictions there are multiple approaches to estimation, for example,
via generalized method of moments. We shall take a more pragmatic approach where we estimate
the parameters of equations separately and then compute
\[
 \frac{1}{2} \hat {\mathsf{a}}'  + \frac{1}{2}  ( {\bcolor \hat \alpha '}  {\pcolor \hat \beta' }+{\pcolor\hat \pi'} ) \text{ and }
 \frac{1}{2} \hat{\mathsf{b}}'  + \frac{1}{2}  ( {\bcolor \hat \alpha '}  {\icolor \hat \gamma'} + {\icolor \hat \mu'}),
\]
as the estimator of the total policy effect. Under standard regularity conditions,
these estimators concentrate around  their population analogues
\[
 \frac{1}{2}  {\mathsf{a}}'  + \frac{1}{2}  ( {\bcolor \alpha '}  {\pcolor \beta' }+{\pcolor \pi'} ) \text{ and }
 \frac{1}{2} {\mathsf{b}}'  + \frac{1}{2}  ( {\bcolor  \alpha '}  {\icolor \gamma'} + {\icolor \mu'}),
\]
with approximate deviations controlled by the normal laws, with standard deviations that can be approximated by the bootsrap
resampling of observational units $i$. Under correct specification the target quantities
reduce to $$  {\bcolor \alpha '}  {\pcolor  \beta' }+{\pcolor \pi'} \text{ and } {\bcolor  \alpha '}  {\icolor  \gamma'} + {\icolor  \mu'},$$ respectively. This construction is not as efficient as generalized method of moments
but has a nicer interpretation under possible misspecification of the model: we are combining predictions
from two models, one motivated via the causal path, reflecting contextual knowledge, and another from a ``reduced form" model not exploiting the path. The combined estimator can improve on precision of either estimator.

>>>>

Our estimates imply that nationally implementing mandatory face masks
for employees in public businesses on March 14th would have reduced the
growth rate of cases and that of deaths by approximately 10 percentage
points in late April. As shown in Figure \ref{fig:US-mask-dgrowth},
this leads to reductions of $21$\% and $34$\% in cumulative reported
cases and deaths, respectively, by the end of May with 90 percent
confidence intervals of $[9,32]$\% and $[19,47]$\%, which roughly
implies that 34 thousand lives could have been saved.
%\footnote{These estimates
%are slightly lower than those reported in earlier releases of the paper on ArXiv.
%One of the reason is that we chose to focus on a simplified version of the model 
%as our base model, by averaging business closure policies as measure of shutdown. This produced more stable results when
%conducting sensitivity results, in particular using fixed effects estimators.  
%We also present several variants of base model, including
%one with separate shutdown policies, in which the effects are slightly larger, in line
%with the previous findings, but we focus reporting counterfactuals only 
%for the base model. } 
This finding is
significant: given this potentially large benefit of reducing the
spread of Covid-19, mandating masks is an attractive policy instrument
especially because it involves relatively little economic
disruption. These estimates contribute to the ongoing efforts towards
designing approaches to minimize risks from reopening
\citep{stock2020b}.

@@

With exclusion restrictions there are multiple approaches to estimation, for example,
via generalized method of moments. We shall take a more pragmatic approach where we estimate
the parameters of equations separately and then compute
\[
 \frac{1}{2} \hat {\mathsf{a}}'  + \frac{1}{2}  ( {\bcolor \hat \alpha '}  {\pcolor \hat \beta' }+{\pcolor\hat \pi'} ) \text{ and }
 \frac{1}{2} \hat{\mathsf{b}}'  + \frac{1}{2}  ( {\bcolor \hat \alpha '}  {\icolor \hat \gamma'} + {\icolor \hat \mu'}),
\]
as the estimator of the total policy effect. Under standard regularity conditions,
these estimators concentrate around  their population analogues
\[
 \frac{1}{2}  {\mathsf{a}}'  + \frac{1}{2}  ( {\bcolor \alpha '}  {\pcolor \beta' }+{\pcolor \pi'} ) \text{ and }
 \frac{1}{2} {\mathsf{b}}'  + \frac{1}{2}  ( {\bcolor  \alpha '}  {\icolor \gamma'} + {\icolor \mu'}),
\]
with approximate deviations controlled by the normal laws, with standard deviations that can be approximated by the bootstrap
resampling of observational units $i$. Under correct specification the target quantities
reduce to $$  {\bcolor \alpha '}  {\pcolor  \beta' }+{\pcolor \pi'} \text{ and } {\bcolor  \alpha '}  {\icolor  \gamma'} + {\icolor  \mu'},$$ respectively.\footnote{This construction is not as efficient as generalized method of moments
but has a nicer interpretation under possible misspecification of the model: we are combining predictions
from two models, one motivated via the causal path, reflecting contextual knowledge, and another from a ``reduced form" model not exploiting the path. The combined estimator can improve on precision of either estimator.}
